Quantitative Benchmarking for Automatically Generated Conjectures

### 2018-03-09 TEBenchmarkPaper/article.tex
Prose

The following issues had to be overcome for the benchmarking to be successful:

\begin{itemize}
  polyrec can't be represented
  constructors were being included
  proofs had to be skipped
  partial functions needed adding to the constraints
\end{itemize}

@@

The most striking result is how rapidly IsaCoSy's running time increases with
sample size: from size 8 upwards, all runs failed (most timed out, a few aborted
early due to running out of memory).

@@

We compare the performance of QuickSpec and IsaCoSy using a similar protocol to
the Speedup-Test of Touati, Worms and Briais~\cite{touati2013speedup}. We differ
by timing a wider variety of inputs: running many samples once, rather than a
few samples many times and taking the medians. Since each sample is run through
both tools, we can also gain statistical power by using the \emph{paired} form
of the Wilcoxon signed-rank test.

>>>>

The following issues had to be overcome to make our benchmark applicable to
IsaCoSy:

\begin{itemize}
\item TIP includes a benchmark called \texttt{polyrec} whose types cannot be
  encoded in Isabelle. We strip out this type and the functions which depend on
  it before translating.
\item When using a type in an IsaCoSy signature, that type's constructors will
  automatically be included in the exploration. Since those constructors will
  not appear in the ground truth (we use $\eta$-expanded wrappers instead, and
  even those may not be present in the current sample) this will unfairly reduce
  the tool's precision. To avoid this, we add a post-processing step which
  replaces all occurrences of a constructor with the corresponding wrapper, then
  discards any conjectures which involve functions other than those in the
  current sample.
\item Since Isabelle is designed for theorem proving rather than programming, it
  requires every definition to be accompanied by proofs of exhaustiveness and
  termination. These are difficult to generate automatically, and don't exist in
  the case of partial functions like destructor wrappers. Hence we use the
  ``quick and dirty'' option in Isabelle, which lets us skip these proofs with
  the \texttt{sorry} keyword.
\item Partial functions cause problems during exploration, since they can throw
  an ``undefined'' exception which causes IsaCoSy to abort. We avoid this by
  pre-populating IsaCoSy's constraint set with these undefined expressions
  (for example \texttt{head Nil}), hence preventing IsaCoSy from ever generating
  them.
\end{itemize}

@@

The most striking result is how rapidly IsaCoSy's running time increases with
sample size: from size 8 upwards, all runs failed (most timed out, a few aborted
early due to running out of memory). As we found with our preliminary testing,
this increase appears exponential, so even large increases to the timeout would
not produce many more successes.

@@

We compare the running times of QuickSpec and IsaCoSy using a similar protocol
to the Speedup-Test of Touati, Worms and Briais~\cite{touati2013speedup}. We
differ by timing a wider variety of inputs: running many samples once, rather
than a few samples many times and taking the medians. Since each sample is run
through both tools, we can also gain statistical power by using the
\emph{paired} form of the Wilcoxon signed-rank test. We compare each sample size
is compared separately, to find the probability that QuickSpec is significantly
faster.
